Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-08-02 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Move creation of all storms and all FIPS to set-up

data(closest_dist)
data(rain)
```

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
# Vector of all storms in study.

storms <- unique(closest_dist$storm_id)
storms <- storms[which(gsub("*.+-", "", storms) <= 2011)]
```

Our analysis included `r length(storms)` Atlantic basin tropical storms over the years 1988 to 2011. This dataset includes several noteworthy storms, including Hugo (1989), Andrew (1992), Floyd (1999), Katrina (2005), [maybe add Ike, Ivan, others?] ... .

```{r echo = FALSE, message = FALSE, warning = FALSE}
# Map of tracks of all storms in study.
map_tracks(storms, plot_points = FALSE, alpha = 0.3)
```

@@

[Spline vs. linear interpolation? Is 15 minutes adequate? Should we use shapefiles instead? IBTrACS probably has this (`rnoaa`).]

@@

To explore if this window (two days before to one day after the storm's closest approach) was adequately large, we investigated patterns in rainfall by day (from four days before to two days after the storm's closest approach) for a few storms that had the potential to not fit this pattern, Floyd (1999), Irene (2011), Lee (2011), Ike (2008), and Allison (2001) (results shown in "validate_precip.pdf").

@@

```{r warning = FALSE, message = FALSE, echo = FALSE}
# Generate a list of unique study fips (all in the eastern US)
my_fips <- closest_dist$fips
my_fips <- unique(my_fips)
storm_lag <- county_rain(counties = my_fips,
                             start_year = 1988, end_year = 2011,
                             rain_limit = 150, dist_limit = 500,
                             days_include = -5:3)
storm_lag <- storm_lag %>%
        mutate(storm_county = paste0(storm_id, "_", fips))

@@

```{r echo = FALSE}
# Generate a list of unique study fips (all in the eastern US)
my_fips <- closest_dist$fips
my_fips <- unique(my_fips)

all_storm_counties <- closest_dist %>%
        dplyr::select(storm_id, fips) %>%
        dplyr::mutate(storm_county = paste(storm_id, fips, sep = "-")) %>%
        dplyr::filter(as.numeric(gsub(".*-", "", storm_id)) >= 1996 & 
                              as.numeric(gsub(".*-", "", storm_id)) <= 2011)

@@

```{r echo = FALSE, warning=FALSE, message=FALSE}
# Generate a list of unique study fips (all in the eastern US)
my_fips <- closest_dist$fips
my_fips <- unique(my_fips)

# Get distance-based exposure for all study counties, all years, using a 
# distance limit that will ensure at least 1000 storm-county pairs
distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
                                     end_year = 2011, dist_limit = 100)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_counties <- unique(closest_dist$fips)

all_distance <- county_distance(counties = all_counties, start_year = 1988, 
                        end_year = 2011, dist_limit =100)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_counties <- unique(closest_dist$fips)

all_rain <- county_rain(counties = all_counties, start_year = 1988, 
                        end_year = 2011, rain_limit = 75, dist_limit = 500,
                        days_included = -1:1)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_counties <- unique(closest_dist$fips)

all_wind <- county_wind(counties = all_counties, start_year = 1988, 
                        end_year = 2011, wind_limit = 18)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_counties <- unique(closest_dist$fips)

all_wind <- county_wind(counties = all_counties, start_year = 1988, 
                        end_year = 2011, wind_limit = 33)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_counties <- unique(closest_dist$fips)

all_wind <- county_wind(counties = all_counties, start_year = 1988, 
                        end_year = 2011, wind_limit = 50)

>>>>

data(closest_dist)
data(rain)

# Generate a list of unique study fips (all in the eastern US)
my_fips <- unique(closest_dist$fips)

# Vector of all storms in study.
storms <- unique(closest_dist$storm_id)
storms <- storms[gsub("*.+-", "", storms) <= 2011 & 
                        gsub("*.+-", "", storms) >= 1988]
```

@@

Our analysis included `r length(storms)` Atlantic basin tropical storms over the years 1988 to 2011. This dataset covers several noteworthy storms, including Hugo (1989), Bob (1991), Andrew (1992), Opal (1995), Fran (1996), Georges (1998), Floyd (1999), Allison (2001), Isabel (2003), Charley (2004), Frances (2004), Ivan (2004), Jeanne (2004), Dennis (2005), Katrina (2005), Rita (2005), Gustav (2008), Ike (2008), and Irene (2011), all of which were severe enough their names were retired.

```{r echo = FALSE, fig.align = 'center', fig.cap = "Tracks of tropical cyclones included in this study. Darker tracks indicate storms whose names were retired."}
# Map of tracks of all storms in study.
a <- map_tracks(storms, plot_points = FALSE, alpha = 0.3, color = "darkcyan")
# Add storms whose names were retired
# Source: http://www.nhc.noaa.gov/aboutnames_history.shtml
map_tracks(c("Hugo-1989", "Bob-1991", "Andrew-1992", 
             "Opal-1995", "Fran-1996", "Georges-1998", "Mitch-1998",
             "Floyd-1999", "Allison-2001", "Michelle-2001",
             "Isidore-2002", "Lili-2002", "Isabel-2003",
             "Charley-2004", "Frances-2004",
             "Ivan-2004", "Jeanne-2004", 
             "Dennis-2005", "Katrina-2005", "Rita-2005",
             "Wilma-2005", "Noel-2007",
             "Gustav-2008", "Ike-2008", "Paloma-2008",
             "Irene-2011"),
           plot_object = a, plot_points = FALSE, color = "darkcyan") + 
        theme(panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           panel.background = element_rect(colour = "black", size = 1))
```

@@

[Is 15 minutes adequate? We could test if we get different results using a finer time resolution.]

@@

```{r warning = FALSE, message = FALSE, echo = FALSE}
storm_lag <- county_rain(counties = my_fips,
                             start_year = 1988, end_year = 2011,
                             rain_limit = 150, dist_limit = 500,
                             days_include = -5:3)
storm_lag <- storm_lag %>%
        mutate(storm_county = paste0(storm_id, "_", fips))

To explore if this window (two days before to one day after the storm's closest approach) was adequately large, we also investigated patterns in rainfall by day (from four days before to two days after the storm's closest approach) for a few storms that had the potential to not fit this pattern, Floyd (1999), Irene (2011), Lee (2011), Ike (2008), and Allison (2001) (results shown in "validate_precip.pdf").

@@

@@

```{r echo = FALSE}
all_storm_counties <- closest_dist %>%
        dplyr::select(storm_id, fips) %>%
        dplyr::mutate(storm_county = paste(storm_id, fips, sep = "-")) %>%
        dplyr::filter(as.numeric(gsub(".*-", "", storm_id)) >= 1996 & 
                              as.numeric(gsub(".*-", "", storm_id)) <= 2011)

@@

```{r echo = FALSE, warning=FALSE, message=FALSE}
# Get distance-based exposure for all study counties, all years, using a 
# distance limit that will ensure at least 1000 storm-county pairs
distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
                                     end_year = 2011, dist_limit = 100)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_distance <- county_distance(counties = my_fips, start_year = 1988, 
                        end_year = 2011, dist_limit =100)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_rain <- county_rain(counties = my_fips, start_year = 1988, 
                        end_year = 2011, rain_limit = 75, dist_limit = 500,
                        days_included = -1:1)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_wind <- county_wind(counties = my_fips, start_year = 1988, 
                        end_year = 2011, wind_limit = 18)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_wind <- county_wind(counties = my_fips, start_year = 1988, 
                        end_year = 2011, wind_limit = 33)

@@

```{r echo = FALSE, message = FALSE, warning = FALSE}
all_wind <- county_wind(counties = my_fips, start_year = 1988, 
                        end_year = 2011, wind_limit = 50)
